---
title: "asset analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#library(HousingVsRenting)
devtools::load_all()

loan_specs <- tibble::tribble(
  ~ loan_years, ~ int_rate_pct,
    30,             4.974,
    20,             4.55,
    15,             4.142
  )

purchase_price <- 1235000

property_tax_pct <- 0.93

insurance_monthly <- 197 # home insurance per month

# Simulated appreciation per year (eg. 0.03 is 3% increase year over year)
appreciation_home_annual <- 0.03
```


```{r build_models, echo=FALSE, include=FALSE}
# Build mortgage payment objects for each down payment
# -- the down payments are hard coded as 200K to 400K
#       in intervals of $50k.
mortg_by_downpay <- function(purchase_price, int_rate_pct, loan_years){
  foo <- function(d) {kahnModel(purchasePrice = purchase_price, downpayment = d,
                                
                                interestRate = int_rate_pct/100, 
                                principalAmortizaion = loan_years,
                                compound_periods = 12)}
  # mortgage payment objects across down payments
  Map(foo, d = list((10^5)*2, (10^5)*2.5, (10^5)*3, (10^5)*3.5, (10^5)*4))
  #Map(foo, d = list((10^5)*3, (10^5)*3.5, (10^5)*4, (10^5)*4.5, (10^5)*5))
}

format_downpayment_cols <- function(x){
  names(x) <- c("200k", "250k", "300k", "350k", "400k")
  #names(x) <- c("300k", "350k", "400k", "450k", "500k")
  
  format(round(x, 0), nsmall=0, big.mark=",") 
}

# Extract monthly mortgage payment from mortgage 
#   payment objects for each down payment.
# -- the down payments are hard coded as 200K to 400K
#       in intervals of $50k.
get_pay_monthly <- function(mortg_){
    pay_monthly <- unlist(Map(function(x) x[["fixedMortgagePayment"]], x = mortg_))
    
  format_downpayment_cols(pay_monthly)
}

mortg_by_downpay_all <- purrr::pmap(.l = loan_specs, 
                                    .f = mortg_by_downpay,
                                    purchase_price = purchase_price)

pay_monthly_all <- purrr::map(mortg_by_downpay_all, .f = get_pay_monthly) %>%
    dplyr::bind_rows() %>%
    dplyr::mutate('Loan years' = loan_specs$loan_years) %>%
  dplyr::relocate('Loan years')

##### Lifetime Interest ####
lifetime_int_all <-  purrr::map(.x = mortg_by_downpay_all, 
                                    .f = function(mortg_by_downpay){
  purrr::map_dbl(.x = mortg_by_downpay,
                 .f = function(mortg_){
                   with(data = mortg_,
                        (fixedMortgagePayment * 12 * principalAmortizaion) - (purchasePrice - downpayment) )
                   }) %>%
                                        format_downpayment_cols(.)
                                      }) %>%
    dplyr::bind_rows() %>%
    dplyr::mutate('Loan years' = loan_specs$loan_years) %>%
  dplyr::relocate('Loan years') 

##### Simulate home appreciation   ####

# home value models by amortization-down payment
home_Value_by_downpay_all <- purrr::map(.x = mortg_by_downpay_all, 
                                    .f = function(mortg_by_downpay){
                                      purrr::map(.x = mortg_by_downpay,
                 .f = function(mortg_){

  homeValue(mortg_,
           propertyTaxRate = property_tax_pct / 100, 
           maintenance = 2500,
           housingAssociationDues = 0, 
           insurance = insurance_monthly * 12,
           appreciationHome = appreciation_home_annual, 
           marginalIncomeTaxRate = 0.26,
           inflation = 0.02,
           time_years = 20)
                   }) 
                                    })
Home_Value_df <- purrr::map(home_Value_by_downpay_all,
                            .f = function(homeValue_by_downpay){
                                      purrr::map(.x = homeValue_by_downpay,
                 .f = function(homeValue_){ 
                   homeValue_$get_valueDF() %>%
                     dplyr::mutate(
                       downpayment = homeValue_[["downpayment"]],
                       loan_years = homeValue_[["principalAmortizaion"]]
                                   )
                   }) %>% dplyr::bind_rows()
                              }) %>% dplyr::bind_rows()
                                    

#####    ####



```

cash: $410k


# cash-flow

What the amount we want to put into a downpayment?

For a $ `r purchase_price/1000` K house the monthly mortgage payment (no fees nor taxes) 
```{r monyhly_pay, echo=FALSE}
  pay_monthly_all  %>%
  kableExtra::kable(.,
                    caption = "Monthly mortgage payment") %>%
  kableExtra::add_header_above(header = c(" " = 1, "Down Payment" = 5)) %>%
  kableExtra::kable_styling()  %>%
  kableExtra::footnote(general = c("Fees & taxes not included"))
```



```{r, echo=FALSE}

tibble::tribble(
  ~'', ~`monthly cost`, ~`yearly cost`, 
  'property tax', 
  purchase_price * property_tax_pct / 100 / 12,
  purchase_price * property_tax_pct / 100,
  'Home Insurance', insurance_monthly, insurance_monthly * 12) %>%
  kableExtra::kable(.,
                    caption = "Monthly costs:") %>%
  kableExtra::kable_styling(full_width = FALSE, position = "left")
```



```{r, echo=FALSE}
lifetime_int_all %>%
  kableExtra::kable(.,
                    caption = "Lifetime interest paid:") %>%
  kableExtra::add_header_above(header = c(" " = 1, "Down Payment" = 5)) %>%
  kableExtra::kable_styling()

```




```{r, echo=FALSE}
year_max <- 10

tmp_ <- Home_Value_df %>%
  dplyr::filter(Time <= year_max * 12)

tmp_ %>%
  dplyr::mutate(
    #time_months = lubridate::duration(Time, units = "months"),
    #Time = lubridate::as.period(time_months, unit = "years"),
    Time = Time / 12,
    downpayment = paste0("$",
                         as.character(downpayment/1000),       
                         " k") ) %>%
  ggplot2::ggplot(data = ., ggplot2::aes(x = Time, y = salesProfit,
                                         color = downpayment)) +
  ggplot2::geom_point() +
  ggplot2::facet_wrap(~loan_years, ncol = 3) +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::scale_x_continuous(breaks = (1:year_max)) +
  ggplot2::labs(x = "\n Time (Years) ", 
                y = "Net Profit \n", 
                title = "Net Profit \n from home sale \n",
                caption = "") + 
 ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5), 
                axis.title.x =  ggplot2::element_text(face="bold", 
                                                      colour="#FF7A33", 
                                                      size = 12),
                axis.title.y =  ggplot2::element_text(face="bold", 
                                                      colour="#FF7A33",
                                                      size = 12),
                legend.title =  ggplot2::element_text(face="bold", 
                                                      size = 10))

```

```{r}
format_dollar_col <- function(x){
   paste0("$",
          format(round(x/1000, 0) * 1000, 
                 nsmall =0 , big.mark = ","))
}

tmp_ %>%
      dplyr::mutate(
     closing_cost = format_dollar_col(sale_transaction_cost(propertyValue)),
     cap_tax = format_dollar_col(capitol_gains_tax(basis = purchase_price, 
                       adj_sales = propertyValue - sale_transaction_cost(propertyValue), 
                       tax_rate = 0.20, 
                       tax_limit = 5*10^5) ),
             propertyValue = format_dollar_col(propertyValue),
        downpayment = paste0("$",
       format(round(downpayment/1000, 0), 
              nsmall =0 , big.mark = ","), " k") ) %>%
  dplyr::group_by(loan_years, downpayment) %>%
  dplyr::mutate(
    housingCosts = format(round(cumsum(housingCosts)/1000,
                                0) * 1000, 
                          nsmall=0, big.mark=",") 
  ) %>%
  dplyr::ungroup() %>%
  dplyr::filter(Time == year_max * 12) %>%
  dplyr::select(propertyValue, closing_cost, cap_tax,
                downpayment, loan_years, housingCosts) %>%
  dplyr::rename(
    "Home Sale Value" = propertyValue, 
    "Loan length (years)" = loan_years,
    "Closing Costs<sup>b</sup>" = closing_cost, 
    "Capitol Gains Tax" = cap_tax
  ) %>%
  unique() %>%
  tidyr::pivot_wider(names_from = downpayment,
                     names_prefix = "",
                     values_from = housingCosts) %>%
  kableExtra::kable(.,
                    caption = sprintf("Statistics for sale at year %s",
                                      year_max),
                    escape = FALSE) %>%
  kableExtra::add_header_above(header = c(" " = 4,
                   "Down Payments" = 5)) %>%
  kableExtra::add_header_above(header = c(" " = 3,
                   "Cumulative Home Costs<sup>a</sup>" = 6),
                    #"Cumulative Home Costs<sup>a</sup>" = 5),
                   #paste0("Cumulative Home Costs", kableExtra::footnote_marker_alphabet(1)) = 5),
                 escape = FALSE) %>%
  kableExtra::kable_styling() %>%
  kableExtra::collapse_rows(columns = 1:3) %>%
  kableExtra::footnote(alphabet = c("Home costs include: inital home purchase costs, interest on debt, home insurance, HOA, maintenance, property tax, & interest deduction", "Closing costs: Broker's fee of 6% & Real Estate Excise 1.28% of property sale"))
```

# liquidity (cash on hand for crisis or home improvement)

- Medical emergency expense
- Car issues
- House Improvements (urgent)
- closing costs
- Move in fees
- Remodeling (construction to aesthetics)
- living expense

## short term cash reserves


### Costs for closing <sup>\*</sup>
- $2100 Title Insurance fees
- Government recording fees
- \$1400 Escrow company fees (buyer unusually pays \$700)
- $9600 [Real Estate Excise Tax](https://dor.wa.gov/taxes-rates/other-taxes/real-estate-excise-tax) (1.28%  sale price, usually paid by the seller)
- Real Estate commissions
- Document preparation fees
- Mortgage origination fees
- Mortgage lender underwriting fee (maybe same as above) \$1,195
- Home appraisal fees
- Mortgage discount points or prepaid interest
- Credit report processing fees
- Processor fees
- Survey fees
- Home owners insurance (first year upfront)

\* Caliber lists total closing cost range \$4,250 - \$4,750.

**Note**: Waive the appraisal (useless)

# risk-mitigation / diversification

Cash reserved can be invested into other areas
