---
title: "Oil Tank Search"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("magrittr")

storage_tank <- readr::read_csv(file.path("..","data-raw", 
                                          "Underground_Storage_Tank.csv"))

street_address <- "11521 39th AVE NE"

street_number <- stringr::str_split(street_address, " " )[[1]][1]

storage_tank_subset <- storage_tank %>%
  dplyr::filter(grepl(pattern = paste0(#"^", 
                                       street_number),
                      x = Address))
```

[Corddry oil tank decommission query](https://data.seattle.gov/Public-Safety/Corddry-oil-tank-decommission-query/nkmp-yxzx)

Records of Seattle Fire Department (SFD) permits related to decommissioning of a residential heating oil tank, permit code 6103. A record with incomplete tank info indicates that the required follow-up report has not been received by SFD. 
Please note that SFD records begin in 1996 when state requirement was introduced. Decommissioning of a residential heating oil tank might have occurred prior to 1996, in which SFD will not have a record.

(Alternative, not used for reprot) [Underground Storage Tank records](https://data.seattle.gov/Public-Safety/Underground-Storage-Tank-UST-Records-Residential/xvj2-ai6y)

## List of oil tanks

```{r roll30Summary, include=TRUE, echo=FALSE}
DT::datatable(storage_tank_subset, 
                # colnames = c('N days' = 'N_records',
                #              "Std. Dev." = "sd"),
                caption = "City of Seattle Storage Tanks",
                options = list(pageLength = 25))
```
