---
title: "Real Estate Broker Exam Prep"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Real Estate Broker Exam Prep}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(HousingVsRenting)
```

Notes on the Peterson’s Master the Real Estate License Exams.

# State Specific

Washington 
Real Estate Commission 
Department of Licensing 
2000 Fourth Avenue West 
Olympia, WA 98502 
P.O. Box 9015 
Olympia, WA 98507-9015 
Phone: 360-664-6500 
Fax: 360-586-0998 
Web site: www.dol.wa.gov/business/realestate/realestatecommission.html 

## Washington Real Estate Law Rules

https://www.dol.wa.gov/business/realestate/lawsrules.html

Includes:
  The Law Relating to Real Estate Brokers and Managing Brokers

## Washington Approved Courses

Posted Curriculum:
https://www.dol.wa.gov/business/realestate/curriculum.html

Online search:
https://professions.dol.wa.gov/s/course-search


# Law of Agency

**AGENT** 